Open-Set Domain Adaptation through Self-Supervision

### 2022-01-22 Report/Paper_OpenSet.tex

This study investigates the use of a simplified ROS method for object classification on the \textit{Office-Home} dataset \cite{OfficeHome}. Alternative self-supervised tasks as well as the inclusion of center-loss are also considered and their performance on the object classification task is evaluated.

@@

\textbf{Rotation-based Open Set} (ROS) is a specific technique developed by Bucci \etal \cite{Bucci2020}. ROS is a two-stage method for open-set domain adaptation. The first stage separates samples in the target domain into known and unknown categories by training the model on a multi-rotation recognition task. The rotation recognition task includes the use of the center-loss to improve performance by learning a center of the features and minimizing the distances between features and their corresponding centers \cite{Wen2016}. The second stage conducts domain alignment, training both semantic and rotation classifiers to classify known target samples. Bucci \etal \cite{Bucci2020} also propose the use of the harmonic mean of the average class accuracies  for the known and unknown classes as a more robust and balanced evaluation metric.

@@

To solve the open-set domain adaptation problem a simplified version of the ROS method (Figure \ref{fig:ROS}) is used. This version uses a single-head rotation classifier and does not include the center-loss, however the effect of the center-loss is evaluated subsequently as a variation to this method. Alternative self-supervised tasks are also considered, including horizontal flipping and through solving jigsaw puzzles. These variations are analyzed in Section \ref{sec:variations}.

@@

\begin{table}[!htb]
  \centering
  \begin{tabular}{@{}|ll|ccc|@{}}
    \hline
\multicolumn{2}{|c}{\textbf{Domain}} & & & \\
\textbf{Source} & \textbf{Target} & \textbf{OS$^*$} & \textbf{UNK} & \textbf{HOS} \\ \hline
Art & Clipart & 35.42 & 58.02 & 43.99 \\
& Product & & & \\
 & Real World & & & \\ \hline
Product & Clipart & & & \\
& Art & & & \\
& Real World & & & \\ \hline
Real World & Clipart & & & \\
 & Art & & & \\
 & Product & & & \\ \hline
Clipart & Real World & & & \\
 & Art & & & \\
& Product & & & \\
    \hline
  \end{tabular}
  \caption{Accuracy (\%) for the simplified ROS model on the Office-Home dataset}
  \label{tab:results}
\end{table}

@@

Two different variations to the simplified ROS method were also analysed. The first considered the use of alternative self-supervised tasks. The second considered the addition of the center-loss to the rotation recognition task.

@@

\begin{figure}[!htb]
  \centering
   %\includegraphics[width=0.95\linewidth]{Figures/Jigsaw.png}
   \caption{Horizontal flipping of an image.}
   \label{fig:horiz}
\end{figure}

@@

\subsection{Center-Loss}

>>>>

This study investigates the use of a simplified ROS method for object classification on the \textit{Office-Home} dataset \cite{OfficeHome}. Alternative self-supervised tasks as well as the inclusion of center loss are also considered and their performance on the object classification task is evaluated.

@@

\textbf{Rotation-based Open Set} (ROS) is a specific technique developed by Bucci \etal \cite{Bucci2020}. ROS is a two-stage method for open-set domain adaptation. The first stage separates samples in the target domain into known and unknown categories by training the model on a multi-rotation recognition task. The rotation recognition task includes the use of the center loss to improve performance by learning a center of the features and minimizing the distances between features and their corresponding centers \cite{Wen2016}. The second stage conducts domain alignment, training both semantic and rotation classifiers to classify known target samples. Bucci \etal \cite{Bucci2020} also propose the use of the harmonic mean of the average class accuracies  for the known and unknown classes as a more robust and balanced evaluation metric.

@@

To solve the open-set domain adaptation problem a simplified version of the ROS method (Figure \ref{fig:ROS}) is used. This version uses a single-head rotation classifier and does not include the center loss, however the effect of the center loss is evaluated subsequently as a variation to this method. Alternative self-supervised tasks are also considered, including horizontal flipping and through solving jigsaw puzzles. These variations are analyzed in Section \ref{sec:variations}.

@@

\begin{table}[!htb]
  \centering
  \begin{tabular}{@{}|ll|ccc|@{}}
    \hline
\multicolumn{2}{|c}{\textbf{Domain}} & & & \\
\textbf{Source} & \textbf{Target} & \textbf{OS$^*$} & \textbf{UNK} & \textbf{HOS} \\ \hline
Art & Clipart & 35.42 & 58.02 & 43.99 \\
& Product & 42.27 & 46.21 & 44.15 \\
 & Real World & 42.50 & 77.01 & 54.77 \\ \hline
Product & Clipart & 38.05 & 55.94 & 45.29 \\
& Art & 33.53 & 46.24 & 38.87 \\
& Real World & & & \\ \hline
Real World & Clipart & & & \\
 & Art & & & \\
 & Product & & & \\ \hline
Clipart & Real World & & & \\
 & Art & & & \\
& Product & & & \\
    \hline
  \end{tabular}
  \caption{Accuracy (\%) for the simplified ROS model on the Office-Home dataset}
  \label{tab:results}
\end{table}

@@

Two different variations to the simplified ROS method were also analysed. The first considered the use of alternative self-supervised tasks. The second considered the addition of the center loss to the rotation recognition task.

@@

\begin{figure}[!htbp]
  \centering
  %\begin{adjustbox}{minipage=\linewidth,scale=0.8}
  \begin{subfigure}[b]{0.4\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{Figures/FlipRaw.png}
    \caption{Raw image}
    \label{fig:short-a}
  \end{subfigure}
  \begin{subfigure}[b]{0.4\columnwidth}
\centering
    \includegraphics[width=\linewidth]{Figures/FlipHoriz.png}
    \caption{Horizontally flipped}
    \label{fig:short-b}
  \end{subfigure}
%\end{adjustbox}
  \caption{Horizontal flipping of an image, adapted from \cite{Xu22019}.}
  \label{fig:centerloss}
\end{figure}

@@

\subsection{Center Loss}

The center loss was introduced by Wen \etal \cite{Wen2016} as an additional objective function for enhancing the discriminative power of learned deep features. It involves learning the center of the deep features for each class, prioritizing features that are close to the center and adding a penalty to the distance between the deep features and their corresponding centers \cite{Wen2016}. The center loss function is optimised alongisde the soft max loss, with the two losses jointly supervised. A hyperparameter is introduced representing the weight of the center loss contribution, and is tuned to balance the two loss functions.The difference between the features learned by a classifier training with the soft max loss function on the MNIST dataset \cite{MNIST}, compared with the inclusion of the center loss during training is shown in Figure \ref{fig:centerloss}. The features learned with the inclusion of center loss can clearly be seen to provide greater discriminative power.

\begin{figure}[!htbp]
  \centering
  %\begin{adjustbox}{minipage=\linewidth,scale=0.8}
  \begin{subfigure}[b]{0.45\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{Figures/SoftMaxLoss.png}
    \caption{Soft Max Loss}
    \label{fig:short-a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.45\columnwidth}
\centering
    \includegraphics[width=\linewidth]{Figures/CenterLoss.png}
    \caption{Center Loss}
    \label{fig:short-b}
  \end{subfigure}
%\end{adjustbox}
  \caption{Comparison of features learned when including the center loss on the MNIST dataset, adapted from \cite{Wen2016}.}
  \label{fig:centerloss}
\end{figure}

An ablation study was performed to select the most appropriate weight of the center loss contribution.
